---
title: "PSTAT 126 Project Step 1"
author: "Hanya Ansari, Carina Yuen, Daren Aguilera"
date: "2023-10-19"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 4,
                      fig.height = 3,
                      fig.align = 'center')
#install.packages("magrittr")
#install.packages("dplyr")   
library(magrittr) 
library(dplyr)
library(faraway)
library(ggplot2)
library(readxl)

# data <- read.csv("C:/Users/Carina W Yuen/Downloads/archive/winequality-red.csv")
data<- read.csv("C:/Users/Daren/Downloads/wine+quality/winequality-red.csv",sep=';')
```

## Name and Source of Data:

#### Wine Quality Based on Physicochemical Tests from UCI Machine Learning Repository

*No Missing Attribute Values*

**Number of Instances:** red wine: 1599

**Number of Variables:** 12 total, 11 continuous, 1 discrete.

### Description of relevant variables and their observational unit:

<!-- ###### This is a multivariate dataset with 11 continuous input variables: -->

<!-- fixed_acidity, volatile_acidity, citric_acid, residual_sugar, chlorides, free_sulfur_dioxide, total_sulfur_dioxide, density, pH, sulphates, alcohol and 1 integer output variable: quality (score between 0 and 10). -->

<!-- Summary statistics and interpretation (use skim) -->

##### Continuous:

-   `fixed acidity` [$\frac{\text{(grams of tartaric acid)}}{dm^3}$]:

    -   acids involved with wine or fixed or nonvolatile (do not evaporate readily)

    ```{r Summary Statistic per variable}
    summary(data$fixed.acidity)
    ```

-   `volatile acidity` [$\frac{\text{grams of acetic acid}}{dm^3}$]:

    -   the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

    ```{r}
    summary(data$volatile.acidity)
    ```

-   `citric acid` [$\frac{g}{dm^3}$]:

    -   found in small quantities, citric acid can add 'freshness' and flavor to wines

    ```{r}
    summary(data$citric.acid)
    ```

-   `residual sugar` [$\frac{g}{dm^3}$]:

    -   the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

    ```{r}
    summary(data$residual.sugar)
    ```

-   `chlorides` [$\frac{\text{grams of NaCl}}{dm^3}$]:

    -   the amount of salt in the wine

    ```{r}
    summary(data$chlorides)
    ```

-   `free sulfur dioxide` [$\frac{mg}{dm^3}$]:

    -   the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

    ```{r}
    summary(data$free.sulfur.dioxide)
    ```

-   `total sulfur dioxide` [$\frac{mg}{dm^3}$]:

    -   amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

    ```{r}
    summary(data$total.sulfur.dioxide)
    ```

-   `density` [$\frac{g}{cm^3}$]:

    -   the density of water is close to that of water depending on the percent alcohol and sugar content

    ```{r}
    summary(data$density)
    ```

-   `pH`:

    -   describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

    ```{r}
    summary(data$pH)
    ```

-   `sulphates` [$\frac{grams_{\text{potassium sulfate}}}{dm^3}$]:

    -   a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

    ```{r}
    summary(data$sulphates)
    ```

-   `alcohol` [$\text{volume percent}$]:

    -   the percent alcohol content of the wine

    ```{r}
    summary(data$alcohol)
    ```

##### Discrete:

-   `Quality`,our output variable, was measured by independent scaling of certified individuals across 3 separate cultivators in Italy. Regarding the preferences, each sample was evaluated by a minimum of three sensory assessors (using blind tastes). 
    It is acknowledged that quality rating via taste is not as clearly defined as continuous quantitative variable.
    -   Observational Unit: Each expert graded the wine quality between 0 (very bad) and 10 (very excellent).

```{r Total summary Statistics}
# data <- read.csv("C:/Users/Carina W Yuen/Downloads/archive/winequality-red.csv")
summary(data)
```

### Graphical Displays and Comments

```{r Fixed acidity plot}
ggplot(data,aes(x=fixed.acidity)) + geom_histogram(bins=30)+  scale_x_continuous(breaks = seq(0, 16, by = 1))+
  ggtitle("Red Wine Fixed Acidity Histogram")
```

The Fixed Acidity histogram suggests that the data may have a gamma distribution, as there is a positive skew, with much of the fixed.acidity values concentrated around 7-8. Comparing the fixed acidity to the volatile acidity, the similar distribution shapes suggest that changing acidity settings does not make a large difference in wine quality in this scenario.

```{r volatile acidity plot}
ggplot(data,aes(x=volatile.acidity)) + 
geom_histogram(bins=30)+
ggtitle("Red Wine Volatile Acidity Histogram")+
scale_x_continuous(breaks = seq(0, 2, by = 0.2))
```

The data for volatile acidity also has a positive skew, with values concentrated around 0.4 and 0.6.

```{r pH plot}
ggplot(data,aes(x=pH)) + geom_histogram(bins=30)+
ggtitle("Red Wine pH Histogram") +
scale_x_continuous(breaks = seq(0, 5, by = 0.1))
```

The pH data has a bell shaped and symmetric distribution, with values centered around 3.3.

```{r quality histogram}
ggplot(data,aes(x=quality)) + geom_histogram(bins=30)+ggtitle("Quality Histogram")
```

The quality data is discrete.

```{r}
ggplot(data,aes(x=citric.acid)) + geom_histogram(bins=30)+ggtitle("Citric Acid Histogram")+scale_x_continuous(breaks = seq(0, 2, by = 0.1))
```

The citric acid data roughly follows a gamma distribution with most of the values concentrated at 0 and 0.475. The presence of two peaks was unexpected.

Linear Model Fit model where y, the outcome variable, is the quality score, and each input variable (acidity, density etc.) is assigned as following.

$$fixed\: acidity =\: x_f, \: volatile\: acidity = x_v,\: citric \: acid\: = x_c,\:  residual \: sugar = x_r,  \: chlorides = x_{ch}, \:free\: sulfur = x_{fs}$$ $$total \: sulfur = x_{ts}, \:  density = x_d,\:  pH = x_p,\:  sulphates = x_s, \:  alcohol = x_a$$

```{r}
cor(data)
```

We did not expect a lot of variation within the data, but upon looking at the summary statistics outputted from skimr, there are certain parameters such as fixed acidity (Min: 4.6 Max: 15.9) and total sulfur dioxide (Min: 6 Max: 289.00) that had a large range compared to the other variables. On the other hand, comparing fixed and volatile acidity, we noticed that variation in acidity levels does not make a large difference in wine quality.

```{r}
 quality_data <- data$quality
 x_f <- data$fixed.acidity
 x_v <- data$volatile.acidity
 x_c <- data$citric.acid
 x_r <- data$residual.sugar
 x_ch <- data$chlorides
 x_fs <- data$free.sulfur.dioxide
 x_ts <- data$total.sulfur.dioxide
 x_d <- data$density
 x_p <- data$pH
 x_s <- data$sulphates
 x_a <- data$alcohol
 fit_quality <- lm(quality_data~x_f+x_v+x_c+x_r+x_ch+x_fs+x_ts+x_d+x_p+x_s+x_a)
 summary(fit_quality)
 
```

The summary statistics indicate that predictor x_s, representing sulphate values, is the predictor with the largest positive effect (and largest effect in magnitude overall) on quality score, with an estimated value of 0.9163. This indicates that this may be a variable of interest and further study to improve wine quality, as there may be a positive relationship between sulphate values and score. On the other hand, the predictor that has the largest negative effect is the density, with an estimated value of -17.88.

Goodness of Fit Analysis

The summary() function outputted two values of the coefficient of determination,with the Multiple R-squared = 0.3606, Adjusted R-squared = 0.3561 . This shows that the quality of the fit is low, as an optimal R-squared value is closer to +1.
